---
title: "Take a taxonomically-aware subset of NRPS clusters for manual testing"
output: html_document
---

```{r}
taxonomy <- read.table("../3_gtdb_taxonomy/tree_taxonomy.tsv",
                        sep='\t', header=T, row.names = 1, colClasses = "character")
```

Only using strains that have at least one NRPS. I've already scanned for NRP metallophores but that info won't be considered during the manual curation.
```{r}
# Get cluster table, which includes all NRPSs found in the representative genomes
all_clusters <- read.table("../2_refseq_reps_results/scores_0625.tsv", sep='\t', header=TRUE)

# Split by strain
per_strain <- split(all_clusters, all_clusters$Assembly)
names(per_strain) <- substr(names(per_strain), 5, 13)

# Only use strains in GTDB
per_strain <- per_strain[rownames(taxonomy)]

# Only use strains with 1+ NRPS
has_nrps <- sapply(per_strain, function(x) sum(x$NRPS == "TRUE") > 0)
```

Dereplicate to one strain per genus
```{r}
has_nrps_taxonomy <- taxonomy[has_nrps, ]

per_genus <- split(has_nrps_taxonomy, has_nrps_taxonomy$G)

test_set <- sapply(per_genus, function(x) {
  rownames(x[sample(nrow(x), 1),])
})

test_set <- paste0("GCF_", test_set)
```

```{r, eval = F}
fileConn<-file("manual_test_set.txt")
writeLines(test_set, fileConn)
close(fileConn)
```

Preparation for scanning with vanilla antiSMASH
```{bash, eval = F}
## Extract all NRPS clusters from the directory of antiSMASH results
for file in */*region*.gbk; do
  if [[ $(basename $file) == *genomic* ]];
    then continue;
  fi;
  grep -l "product=\"NRPS\"" $file >> nrp_regions.txt;
done

for f in $(cat nrp_regions.txt); do
  g=${f//\//_};
  cp $f ../all_nrps_Jun25/$g;
done

## Extract ones from the test set taxa
for f in $(cat nrps_test_set.txt); do
  find ../all_nrps_Jun25 -name "$f*" -exec cp "{}" nrps_Jun25 \;;
done

## Rename headers
for f in *.gbk; do
  sed -i "1s/.*/LOCUS       $f/" $f;
done
```

Genbanks were chunked and run with vanilla antiSMASH 6.1.1.
Results and genbanks are available in `manual_testing_clusters.tar.gz`

From here, see Table S2

```{r}
sessionInfo()
```