---
title: "R Notebook"
output: html_document
---

### Associate taxonomy with the metallophores
```{r}
taxonomy <- read.table("../3_gtdb_taxonomy/tree_taxonomy.tsv",
                       sep='\t', header=TRUE, colClasses = "character")

strain_df <- read.table("../7_by_strain/complete_genome_cluster_counts.tsv",
                        sep='\t', header=TRUE)
rownames(strain_df) <- substr(strain_df$Assembly, 5, 13)

strain_df <- strain_df[taxonomy$X, ]

write.table(cbind(strain_df, taxonomy[, 2:7]),
            "taxonomy_and_metallophores.tsv",
            sep = "\t", quote = F, row.names = F)
```


### Is NRP metallophore biosynthesis unevenly distributed across bacterial taxa?
```{r}
taxonomy$HasNRP.met <- strain_df$NRP.met > 0
taxonomy$HasNRPS <- strain_df$NRPS > 0

phylum_met <- table(taxonomy[,c("P", "HasNRP.met")])

chisq.test(phylum_met)
```

Chi square shouldn't be used for small expected values (rule of thumb is <5). Combining the small phyla:
```{r}
# Overall proportion is 15%, min of 33.3 samples gives >5 expected
phylum_met_cont <- as.data.frame.matrix(phylum_met)
big_enough <- phylum_met_cont[rowSums(phylum_met_cont) > 33.3,]

other <- colSums(phylum_met_cont[rowSums(phylum_met_cont) < 33.3,])
w_other <- rbind(big_enough, other)

chisq.test(w_other)

library(ggplot2)
df <- data.frame(phylum_met)
bigs <- as.character(df$P)
bigs[! bigs %in% rownames(big_enough)] <- "other"

df$P <- as.factor(bigs)
ggplot(df, aes(x = P, y = Freq, fill = HasNRP.met)) +
  geom_bar(position="fill", stat="identity") +
  coord_flip()
```

## Looking at Proteobacteria and Actinobacteria
```{r}
actinos <- taxonomy[taxonomy$P == "Actinobacteriota", ]

orders <- as.data.frame(table(actinos$O))
orders$HasNRP.met <- sapply(orders$Var1, function(taxon) nrow(actinos[actinos$O == taxon & actinos$HasNRP.met,]))
orders$NoNRP.met <- orders$Freq - orders$HasNRP.met
orders$Fraction <- orders$HasNRP.met/orders$Freq
orders[order(orders$Fraction, decreasing = T),]

freq <- orders[, c("HasNRP.met","NoNRP.met")]
rownames(freq) <- orders$Var1

## Combine small taxa to reach expected value >5
min_freq <- 5/(sum(freq$HasNRP.met/sum(freq)))
small <- rowSums(freq) <= min_freq
freq <- rbind(freq[!small, ], colSums(freq[small,]))

freq
chisq.test(freq)

big_orders <- actinos[actinos$O %in% c("Mycobacteriales", "Streptomycetales", "Actinomycetales"), ]

pairwise.t.test(big_orders$HasNRP.met, big_orders$O)
```

```{r}
proteos <- taxonomy[taxonomy$P == "Proteobacteria", ]

orders <- as.data.frame(table(proteos$O))
orders$HasNRP.met <- sapply(orders$Var1, function(taxon) nrow(proteos[proteos$O == taxon & proteos$HasNRP.met,]))
orders$NoNRP.met <- orders$Freq - orders$HasNRP.met
orders$Fraction <- orders$HasNRP.met/orders$Freq
orders[order(orders$Fraction, decreasing = T),]

freq <- orders[, c("HasNRP.met","NoNRP.met")]
rownames(freq) <- orders$Var1

## Combine small taxa to reach expected value >5
min_freq <- 5/(sum(freq$HasNRP.met/sum(freq)))
small <- rowSums(freq) <= min_freq
freq <- rbind(freq[!small, ], colSums(freq[small,]))

freq
chisq.test(freq)
```


## NRP.mets and other NRPSs
```{r}
strains_with_nrps <- taxonomy[strain_df$NRPS > 0, ]

message("Overall fraction")
(frac <- sum(strains_with_nrps$HasNRP.met) / nrow(strains_with_nrps))
min_size <- 5 / frac

phy <- strains_with_nrps[,c("P", "HasNRP.met")]
too_small <- names(table(strains_with_nrps$P))[table(strains_with_nrps$P) < min_size]
phy$P[phy$P %in% too_small] <- "Other"

chisq.test(table(phy))

tab <- as.data.frame.matrix(table(phy))
tab$fraction <- tab$"TRUE" / (rowSums(tab))

tab$pvalue <- sapply(rownames(tab), function(phylum) {
  fisher.test(phy$P == phylum, phy$HasNRP.met)$p.value * length(unique(phy$P))
})
print(data.frame(rownames(tab), tab$fraction, tab$pvalue))
```


```{r}
sessionInfo()
```